### 2014-10-27 core-lang.tex
Fixed typos

We demonstrate our ideas using a replicated data store formalized as a
triple $\E;\Theta;\Sigma$, where $\E$ denotes the execution state,
$\Theta$ maps replicas to their local state, and $\Sigma$, the session
soup, is the set of client sessions interacting with the store.
Operational semantics given in Figure~\ref{sem:oper} succinctly
capture the behaviour of the store that is relevant to demonstrate the
operational significance of contract classification. For the sake of
clarity, we only consider a single replicated object of well-defined
type (for eg: a replicated object of type \cf{BankAccount}) in our
formalization.  Our semantics are parametric over the specification of
this replicated data type.  Figure~\ref{sem:oper} formalizes
replicated data type (RDT) specification as tuple
$(\tau,\Ops,\Ctrts)$, where $\tau$ is the data type, $\Ops$ maps
labels ($op$) of operations on $\tau$ to their definitions, while
$\Ctrts$ maps them to their consistency contracts ($\cv$). The
definition of an operation is expected to be a lambda expression in
the server-side language\footnote{Since a replicated object is
maintained by a server, we call the language in which replicated data
type operations are defined as \emph{server-side language}.}, although
we do not enforce this in our formalization. We intentionally leave
the server-side language underspecified to demonstrate that our ideas
are independent of the choice of the programming language used to
implement the replicated data type.

@@

Clients interact with the store via multiple concurrent sessions. For
technical reasons, we tag every session in the session soup ($\Sigma$)
with a unique {\sf SessID} and the sequence number of next operation
in that session. As shown in Fig.~\ref{sem:oper}, we model a session
simply as a sequence of actions on the replicated object maintained by
the store, where an action is defined as an application of replicated
data type operation to a value. As with the server-side language, our
development is also orthogonal to the choice of the client-side
language, but we require both the languages to agree on the syntactic
class of values that replicated data type operations ($op$) produce
and consume.

An execution state $\E$ is a tuple $\Exec$ where $\EffSoup$ is a set
of effects, and $\visZ,\soZ \subseteq \EffSoup \times
\EffSoup$ are visibility and session order relations over effects,
respectively. An effect ($\eff$), modeled as a tuple $(s,i,op,v)$,
records the fact that $i^{th}$ action in session with {\sf SessID}
$s$, which is an operation $op$ on the replicated object, has been
successfully executed on some replica yielding a return value $v$.
Note that the combination of $s$ and $i$ uniquely identifies the
effect. Session order relation ($\soZ$) relates effects generated by
the same session. An effect $\eff = (s,i,op,v)$ is said to precede
another effect $\eff' = (s',i',op',v')$ in session order if and only
if $s'=s$ and $i'\ge i$. An effect generated at a replica becomes
visible at rest of the replicas eventually. The local state of a
replica $r$ (i.e., $\Theta(r)$) is a set of effects that are currently
visible at $r$. An operation $op$ performed at replica $r$ can only
witness the set of effects ($\Theta(r)$) visible at $r$.  Note that
in presence of multiple replicated objects, $\visZ$ only relates
effects of operations performed on a single object. For instance, a
{\sf withdraw} operation can witness (via $\visZ$) a {\sf deposit}
operation on the same bank account, but not a {\sf deposit} on a
different account. Often, in our formalization, we use $\visZ$ and
$\soZ$ binary relations to obtain a set of effects visible to a given
effect $\eff$, or set of effects that precede a given effect $\eff$ in
the session order. As a syntactic convenience, whenever $R$ is a
binary relation, we write $R(\eff)$ to denote the set of all $\eff'$
such that $(\eff,\eff') \in R$.  Conversely, we write $R^{-1}(\eff)$
to denote the set of all $\eff'$ such that $(\eff',\eff) \in R$.

@@

Rule $\rulelabel{Oper}$ is an auxiliary reduction of the form

\begin{smathpar}
\auxred{\Theta} {(\E,\langle s,i,op(v) \rangle)} {r} {(\E',\eff)}
\end{smathpar}

\noindent Under the store configuration $\Theta$, the rule captures
the progress in execution (from $\E$ to $\E'$) due to the application
of operation $op$ to replica $r$ resulting in a new effect $\eff$.
The rule first constructs a \emph{context} for the application from
the local state ($\Theta(r)$) of the replica, by
projecting\footnote{{\textsf{ctxt*}} is auxiliary function
\textsf{ctxt} extended straightforwardly to set of effects} relevant
information from effects in $\Theta(r)$. It then substitutes the
definition ($\Ops(op)$) of the operation for its label ($op$), and
relies on the reduction relation ($\rdtredsto$) of the server-side
language to reduce the application $\Ops(op)(v,~ctxt)$ to a value
$v'$.  Subsequently, the the attributes of execution state, namely
$\EffSoup$, $\visZ$ and $\soZ$, are extended to incorporate
the new effect ($\eff$).

@@

\begin{definition}[Well-formed execution]
An execution state $\E = \Exec$ corresponding to an execution of
operations over a replicated datatype with specification $\rdtspec =
(\tau,\Ops,\Ctrts)$ is said to be well-formed if and only if 
$\forall \eff \in \EffSoup. ~\E \models \De{\cv[\eff/\cureff]}$, where $\cv =
\Ctrts(\operZ(\eff))$.
\end{definition}

Symbol $\models$ denotes the standard \emph{models} relation of
first-order logic.  A valid model for a first-order formula defines
the universe of discourse, and provides interpretations for free
symbols in the formula such that the formula is true under given
interpretations.  An execution state ($\E$) provides interpretations
for $\Rvis$ and $\Rso$ relations, while fixing the universe of
discourse to be the set of all generated effects ($\EffSoup$).
Intuitively, the well-formedness definition captures the idea that for
every generated effect, the contract on the corresponding operation is
satisfied under the given execution state.

>>>>

We formalize a replicated data store as a triple $\E;\Theta;\Sigma$,
where $\E$ denotes the execution state, $\Theta$ maps replicas to
their local state, and $\Sigma$, the session soup, is the set of
concurrent client sessions interacting with the store.  Operational
semantics given in Figure~\ref{sem:oper} succinctly capture the
behaviour of the store that is relevant to demonstrate the operational
significance of contract classification. For the sake of clarity, we
only consider a single replicated object of well-defined type (for eg:
a replicated object of type \cf{BankAccount}) in our formalization.
Our semantics are parametric over the specification of this replicated
data type.  Figure~\ref{sem:oper} formalizes replicated data type
(RDT) specification as tuple $(\tau,\Ops,\Ctrts)$, where $\tau$ is the
data type, $\Ops$ maps labels ($op$) of operations on $\tau$ to their
definitions, while $\Ctrts$ maps them to their consistency contracts
($\cv$). The definition of an operation is expected to be a lambda
expression in the server-side language\footnote{Since a replicated
object is maintained by a server, we call the language in which
replicated data type operations are defined as \emph{server-side
language}.}, although we do not enforce this in our formalization. We
intentionally leave the server-side language underspecified to
demonstrate that our ideas are independent of the choice of the
programming language used to implement the replicated data type.

@@

A session is modeled simply as a sequence of actions on the replicated
object maintained by the store, where an action is defined as an
application of replicated data type operation to a value. For
technical reasons, we tag every session in the session soup ($\Sigma$)
with a unique {\sf SessID} and the sequence number of next operation
in that session. As with the server-side language, our development is
also orthogonal to the choice of the client-side language, but we
require both the languages to agree on the syntactic class of values
that replicated data type operations ($op$) produce and consume.

An execution state $\E$ is a tuple $\Exec$ where $\EffSoup$ is a set
of effects, and $\visZ,\soZ \subseteq \EffSoup \times
\EffSoup$ are visibility and session order relations over effects,
respectively. An effect ($\eff$), modeled as a tuple $(s,i,op,v)$,
records the fact that $i^{th}$ action in session with {\sf SessID}
$s$, which is an operation $op$ on the replicated object, has been
successfully executed on some replica yielding a return value $v$.
Note that the combination of $s$ and $i$ uniquely identifies the
effect. Session order relation ($\soZ$) relates effects generated by
the same session. An effect $\eff = (s,i,op,v)$ is said to precede
another effect $\eff' = (s',i',op',v')$ in session order if and only
if $s'=s$ and $i'\ge i$. An effect generated at a replica becomes
visible at rest of the replicas eventually. The local state of a
replica $r$ (i.e., $\Theta(r)$) is a set of effects that are currently
visible at $r$. An operation $op$ performed at replica $r$ can only
witness the set of effects ($\Theta(r)$) visible at $r$.  
% Note that in presence of multiple replicated objects, $\visZ$ only
% relates effects of operations performed on a single object. For
% instance, a {\sf withdraw} operation can witness (via $\visZ$) a {\sf
% deposit} operation on the same bank account, but not a {\sf deposit}
% on a different account. 
Often, in our formalization, we use $\visZ$ and
$\soZ$ binary relations to obtain a set of effects visible to a given
effect $\eff$, or set of effects that precede a given effect $\eff$ in
the session order. As a syntactic convenience, whenever $R$ is a
binary relation, we write $R(\eff)$ to denote the set of all $\eff'$
such that $(\eff,\eff') \in R$.  Conversely, we write $R^{-1}(\eff)$
to denote the set of all $\eff'$ such that $(\eff',\eff) \in R$.

@@

Rule $\rulelabel{Oper}$ is an auxiliary reduction of the
form:\vspace{-1.7mm}
\begin{smathpar}
\auxred{\Theta} {(\E,\langle s,i,op(v) \rangle)} {r} {(\E',\eff)}
\vspace{-1.7mm}
\end{smathpar}
\noindent Under the store configuration $\Theta$, the rule captures
the progress in execution (from $\E$ to $\E'$) due to the application
of operation $op$ to replica $r$ resulting in a new effect $\eff$.
The rule first constructs a \emph{context} for the application from
the local state ($\Theta(r)$) of the replica, by
projecting\footnote{{\textsf{ctxt*}} is auxiliary function
\textsf{ctxt} extended straightforwardly to set of effects} relevant
information from effects in $\Theta(r)$. It then substitutes the
definition ($\Ops(op)$) of the operation for its label ($op$), and
relies on the reduction relation ($\rdtredsto$) of the server-side
language to reduce the application $\Ops(op)(v,~ctxt)$ to a value
$v'$.  Subsequently, the the attributes of execution state, namely
$\EffSoup$, $\visZ$ and $\soZ$, are extended to incorporate
the new effect ($\eff$).

@@

\begin{definition}[Well-formed execution]
An execution state $\E = \Exec$ corresponding to an execution of
operations over a replicated datatype with specification $\rdtspec =
(\tau,\Ops,\Ctrts)$ is said to be well-formed if and only if $\forall
\eff \in \EffSoup. ~\E \models \DDe{[\eff/\cureff]\cv}$, where $\cv =
\Ctrts(\operZ(\eff))$.
\end{definition}

$\models$ denotes the standard \emph{models} relation of
first-order logic.  A valid model for a first-order formula defines
the universe of discourse, and provides interpretations for free
symbols in the formula such that the formula is true under given
interpretations.  An execution state ($\E$) provides interpretations
for $\Rvis$ and $\Rso$ relations, while fixing the universe of
discourse to be the set of all generated effects ($\EffSoup$).
Intuitively, the well-formedness definition captures the idea that for
every generated effect, the contract on the corresponding operation is
satisfied under the given execution state.
